---
title: "HEATON_Skills_Test_3.Rmd"
output: html_document
date: "2024-11-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

First the data was tidied by creating a column containing the Rank and corresponding salary of faculty. Once the data was tidied, a graph was created showing the rank 
of different faculty and the corresponding salary. This was also faceted by the tier of
the school.


```{r faculty, warning=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
df <- read.csv('./FacultySalaries_1995.csv')
```
Here is cleaning:
```{r cleaning, echo=TRUE, warning=FALSE, message=FALSE}
cleandf <- df %>%
  pivot_longer(ends_with('Salary'),
             names_to = 'Rank',
             values_to = 'Salary') %>%
  transform(Rank=str_replace(Rank,"Avg","")) %>%
  transform(Rank=str_replace(Rank,"ProfSalary",""))
```
Below is the plot I created:


```{r cleandf, echo=FALSE, warning=FALSE}
cleandf %>%
  filter(Tier != "VIIB") %>%
  ggplot(aes(x = Rank, y = Salary, fill = Rank)) +
  geom_boxplot() +
  facet_grid(~ Tier) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle =45))

```

## Question 2

I created an ANOVA model to see State, Tier, and Rank affect on Salary.
The code used is posted below along with a summary.

```{r creating model}
aov <- cleandf %>% 
  aov(data = .,
      formula = Salary ~ State + Tier + Rank)
summary(aov)
```


## Question 3
Below is the code for tidying the data set Juniper Oils. 

```{r tidying, message=FALSE, results=FALSE}
oil_df <- read.csv('./Juniper_Oils.csv')
clean_names(oil_df)

names(oil_df) <- gsub("\\.", "-", names(oil_df))

new_df <- oil_df %>%
  pivot_longer(cols = c("alpha-pinene","para-cymene","alpha-terpineol","cedr-9-ene",
                        "alpha-cedrene","beta-cedrene","cis-thujopsene","alpha-himachalene","beta-chamigrene",
                        "cuparene","compound-1","alpha-chamigrene","widdrol","cedrol","beta-acorenol","alpha-acorenol",
                        "gamma-eudesmol","beta-eudesmol","alpha-eudesmol","cedr-8-en-13-ol","cedr-8-en-15-ol","compound-2","thujopsenal"),
               names_to = 'ChemicalID',
               values_to = 'Concentration')
```

## Question 4
Then I used the tidied data set to create a graph that showed the relationship between years since burn and the concentration of different chemicals. 

```{r graph}
new_df %>%
  ggplot(aes(x = YearsSinceBurn, y = Concentration)) +
  geom_smooth() +
  facet_wrap(~ ChemicalID, scales = "free_y") +
  theme_minimal()
```

## Question 5
I then used a generalized linear model to see which chemicals had a significant concentration affected by years since burn. 

```{r model and output}
mod1 <- glm(data = new_df,
            formula = Concentration ~ YearsSinceBurn * ChemicalID)

broom::tidy(mod1) %>%
  filter(p.value < 0.05)
```


